---
title: "US Flight Statistics (October 2014)"
output: 
  flexdashboard::flex_dashboard:
    theme: cerulean
    orientation: rows
    vertical_layout: fill
    social: ["twitter", "facebook", "menu"]
---

<style>
.dataTables_scrollBody {
    max-height: 100% !important;
}
</style>

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(DT)
library(rpivotTable)
library(ggplot2)
library(plotly)
library(dplyr)
library(openintro)
library(highcharter)
library(ggmap)
library(maps)
library(geosphere)
```

```{r}
data <- read.csv("US_Airlines")
data$UniqueCarrier <- as.factor(data$UniqueCarrier)
data$Origin <- as.factor(data$Origin)
data$OriginCityName <- as.factor(data$OriginCityName)
data$OriginState <- as.factor(data$OriginState)
data$OriginStateName <- as.factor(data$OriginStateName)
data$Dest <- as.factor(data$Dest)
data$DestCityName <- as.factor(data$DestCityName)
data$DestState <- as.factor(data$DestState)
data$DestStateName <- as.factor(data$DestStateName)
names(data)[names(data) == 'Distan'] <- 'ActualElapsedTime'
data <- data %>% filter(!is.na(ArrDelay))
```

```{r}
mycolors <- c("blue", "FFC125", "darkgreen", "darkorange")
```

Overall Statistics
================================

Row
--------------------------------

### Total Flights
```{r}
valueBox(nrow(data), icon="fa-plane")
```

### Total Carriers
```{r}
valueBox(length(unique(data$UniqueCarrier)), icon="fa-plane")
```

Row
--------------------------------

### Mean Departure Delay in Minutes
```{r}
valueBox(round(mean(data$DepDelay), 2), icon="fa-plane-departure")
```

### Flights with >15 Minutes Departure Delay
```{r}
valueBox(sum(data$DepDel15), icon="fa-plane-departure")
```

Row
--------------------------------

### Mean Arrival Delay in Minutes
```{r}
valueBox(round(mean(data$ArrDelay), 2), icon="fa-plane-arrival")
```

### Flights with >15 Minutes Arrival Delay
```{r}
valueBox(sum(data$ArrDel15), icon="fa-plane-arrival")
```

Row
--------------------------------

### Total Distance Travelled in Miles
```{r}
valueBox(sum(data$Distance), icon="fa-plane")
```

### Total Time Travelled in Minutes
```{r}
valueBox(sum(data$AirTime), icon="fa-plane")
```

Row
--------------------------------

### Total Flights Delayed
```{r}
temp <- data %>% filter(ArrDelayMinutes>0)
valueBox(nrow(temp))
```

### Total Flights Delayed due to Carrier
```{r}
temp <- data %>% filter(CarrierDelay>0)
valueBox(nrow(temp))
```

### Total Flights Delayed due to Weather
```{r}
temp <- data %>% filter(WeatherDelay>0)
valueBox(nrow(temp))
```

### Total Flights Delayed due to Securiry
```{r}
temp <- data %>% filter(SecurityDelay>0)
valueBox(nrow(temp))
```


Tabular Data
================================

```{r}
temp <- data %>% filter(ArrDelay>0) %>% select(FlightDate, UniqueCarrier, FlightNum, OriginCityName, DestCityName, DepTime, ArrTime, DepDelay, ArrDelay)
datatable(temp,
          caption = "Delayed Flights",
          rownames = T,
          filter = "top",
          options = list(pageLength = 10))
```

Carrier Statistics
================================

Row
--------------------------------

### Total Flights Operated by each Carrier

```{r}
temp <- data %>% group_by(UniqueCarrier) %>% summarise(Count = n()) %>% arrange(desc(Count))
v1 <- temp %>% hchart(type='column', color = "coral", hcaes(x = UniqueCarrier, y = Count)) %>% hc_yAxis(title=list(text = "Total Flights")) %>% hc_xAxis(title = list(text = "Carrier"))
v1
```

### Total Flights Delayed by each Carrier

```{r}
temp <- data %>% filter(ArrDelayMinutes>0) %>% group_by(UniqueCarrier) %>% summarise(Count = n()) %>% arrange(desc(Count))
v2 <- temp %>% hchart(type='column', hcaes(x = UniqueCarrier, y = Count)) %>% hc_yAxis(title=list(text = "Total Flights")) %>% hc_xAxis(title = list(text = "Carrier"))
v2
```

Row
--------------------------------

### Average Departure Delay of Flights by each Carrier

```{r}
temp <- data %>% group_by(UniqueCarrier) %>% summarise(Average = round(mean(DepDelay), 2)) %>% arrange(desc(Average))
v3 <- temp %>% hchart(type = 'column', color = 'bisque', hcaes(x = UniqueCarrier, y = Average)) %>% hc_yAxis(title=list(text = "Delay in Minutes")) %>% hc_xAxis(title = list(text = "Carrier"))
v3
```

### Average Arrival Delay of Flights by each Carrier

```{r}
temp <- data %>% group_by(UniqueCarrier) %>% summarise(Average = round(mean(ArrDelay), 2)) %>% arrange(desc(Average))
v4 <- temp %>% hchart(type='column', color='aquamarine', hcaes(x = UniqueCarrier, y = Average)) %>% hc_yAxis(title=list(text = "Delay in Minutes")) %>% hc_xAxis(title = list(text = "Carrier"))
v4
```

State-wise Analysis
================================

### Average Arrival Delay of Flights in each State (Minutes)

```{r}
temp <- data %>% group_by(DestState) %>% summarise(Average = round(mean(ArrDelay), 2)) %>% arrange(desc(Average))
temp$DestState <- abbr2state(temp$DestState)
highchart() %>% hc_add_series_map(usgeojson, temp, name = 'DestState', value = 'Average', joinBy = c('woename', 'DestState')) %>%  hc_mapNavigation(enabled = T)
```

City-wise Analysis
============================================

### Flight Route with >15 Minutes Arrival Delay (Top 15 Cities)

```{r}
# temp <- data %>% filter(ArrDel15 == 1)
# temp <- distinct(temp[c("OriginCityName", "DestCityName")])
# origin_codes <- geocode(as.character(temp$OriginCityName))
# dest_codes <- geocode(as.character(temp$DestCityName))
# names(origin_codes) <- c('lon1', 'lat1')
# names(dest_codes) <- c('lon2', 'lat2')
# temp <- data.frame(temp, origin_codes, dest_codes)
# write.csv(temp,"network1.csv", row.names = FALSE)

temp <- read.csv("network1.csv")
{
map("world", regions=c("usa"), fill=T, col="aliceblue", ylim=c(21.0,50.0), xlim=c(-130.0,-65.0))
points(temp$lon1, temp$lat1, pch=10, cex=0.4, col="coral")
v5 <- distinct(temp %>% group_by(OriginCityName) %>% summarise(Count = n(), lon = lon1, lat = lat1) %>% arrange(desc(Count)))
v5 <- v5[1:15,]
text(v5$OriginCityName, x=v5$lon, y=v5$lat,  col="coral", cex=0.5, pos=1)
for (i in (1:dim(temp)[1])) {
  inter <- gcIntermediate(c(temp$lon1[i], temp$lat1[i]), c(temp$lon2[i], temp$lat2[i]), n=200)
  lines(inter, lwd=0.1, col=alpha("chartreuse", 0.4))
}
}
```

### Flight Route with no Arrival Delay (Top 15 Cities)
```{r}

# temp <- data %>% filter(ArrDelay < 1)
# temp <- distinct(temp[c("OriginCityName", "DestCityName")])
# origin_codes <- geocode(as.character(temp$OriginCityName))
# dest_codes <- geocode(as.character(temp$DestCityName))
# names(origin_codes) <- c('lon1', 'lat1')
# names(dest_codes) <- c('lon2', 'lat2')
# temp <- data.frame(temp, origin_codes, dest_codes)
# write.csv(temp,"network2.csv", row.names = FALSE)

temp <- read.csv("network2.csv")

{
map("world", regions=c("usa"), fill=T, col="aliceblue", ylim=c(21.0,50.0), xlim=c(-130.0,-65.0))
points(temp$lon1, temp$lat1, pch=10, cex=0.4, col="coral")
v5 <- distinct(temp %>% group_by(OriginCityName) %>% summarise(Count = n(), lon = lon1, lat = lat1) %>% arrange(desc(Count)))
v5 <- v5[1:15,]
text(v5$OriginCityName, x=v5$lon, y=v5$lat,  col="coral", cex=0.5, pos=1)
for (i in (1:dim(temp)[1])) {
  inter <- gcIntermediate(c(temp$lon1[i], temp$lat1[i]), c(temp$lon2[i], temp$lat2[i]), n=200)
  lines(inter, lwd=0.1, col=alpha("chartreuse", 0.4))
}
}
```

Summary Statistics
============================================

Row
--------------------------------------------

### Average Arrival Delay for each Day of Month

```{r}
temp <- data %>% group_by(DayofMonth) %>% summarise(Average = mean(ArrDelay)) %>% arrange(desc(Average))
v6 <- temp %>% ggplot(aes(x = DayofMonth, y = Average, group = 1)) + geom_line() + geom_point() + xlab("Day of Month") + ylab("Arrival Delay in Minutes")
v6
```

### Average Arrival Delay for each Hour of Day

```{r}
temp <- data %>% group_by( ArrTime = round(CRSArrTime/100.0) ) %>% summarise(Average = mean(ArrDelay)) %>% arrange(desc(Average))
v7 <- temp %>% filter(ArrTime > 1) %>% ggplot(aes(x = ArrTime, y = Average, group = 1)) + geom_line() + geom_point() + xlab("Hour of Day") + ylab("Arrival Delay in Minutes")
v7
```

Row
--------------------------------------------

### Departure Delay vs Arrival Delay

```{r}
v8 <- data %>% ggplot(aes(x = DepDelay, y = ArrDelay)) + geom_point() + geom_smooth(method = 'lm') + xlab("Departure Delay in Minutes") + ylab("Arrival Delay in Minutes")
v8
```

### Distribution of Flight Delay

```{r}
v9 <- data %>% ggplot(aes(x = ArrDelay)) + geom_histogram(bins=60) + xlim(-60, 120) + xlab("Delay") + ylab("Total Flights")
v9
```
